---
title: "Integrating all the information"
author: "by [Yan Holtz](https://github.com/holtzy/) - `r format(Sys.time(), '%d %B %Y')`"
---


<br><br>
```{r echo=FALSE, warning=FALSE, message=FALSE}
# A few library
library(tidyverse)
library(DT)
library(hrbrthemes)
library(ggrepel)
```

>This document applies the SMR tool on both the Expression and Methylation data in the same time. It also takes into account the descriptive classification of chromatin.






#- Method
***
I'm gonna use the methylation data located on Delta at `/gpfs/gpfs01/polaris/Q0286/uqywu16/omics/methy`

I run SMR on it, like I've done for eQTL data.






#- Run SMR {.tabset}
***

Run SMR on both expression and methylation datasets in the same time.

## Chromo 11 ~ 15Mb
```{r, eval=FALSE}
# Good folder
cd /shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION

# list of option and file:
ref="/gpfs/gpfs01/polaris/Q0286/UKBiobank/v2EURu_HM3/ukbEURu_imp_chr11_v2_HM3_QC"
sum="/shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/1_GWAS/GWAS_vitaminD_XiaEtAL.ma"
METHY="/gpfs/gpfs01/polaris/Q0286/uqywu16/omics/methy/bl_meqtl_cis_std_chr11"
rug="/shares/compbio/Group-Wray/YanHoltz/DATA/EQTL/eQTL_DATA_EQTLGEN_CONSORTIUM_32K/cis-eQTLs-full_formatted.txt_besd-dense"
genelist="/shares/compbio/Group-Wray/YanHoltz/DATA/GENE_POSITION/glist_hg19_refseq.txt"
probe="ENSG00000152270"
window="1000"
psmr="5e-8"
thread="6"
out="./integration_result_Chr11_15Mb"

# Run the analysis
tmp_command="smr_Linux --bfile $ref --gwas-summary $sum --beqtl-summary $METHY  --beqtl-summary $rug --plot --probe $probe --probe-wind $window --gene-list $genelist --psmr $psmr --out $out --thread-num $thread"
qsubshcom "$tmp_command" 1 30G smr_integration 10:00:00 ""

# Transfer locally
cd /Users/y.holtz/Dropbox/QBI/4_UK_BIOBANK_GWAS_PROJECT/VitaminD-GWAS/0_DATA/SMR_PLOT_DATA
scp y.holtz@delta.imb.uq.edu.au:/shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION/plot/* .
```



## Chromo 11 ~ 71Mb
```{r, eval=FALSE}
# Good folder
cd /shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION

# list of option and file:
ref="/gpfs/gpfs01/polaris/Q0286/UKBiobank/v2EURu_HM3/ukbEURu_imp_chr11_v2_HM3_QC"
sum="/shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/1_GWAS/GWAS_vitaminD_XiaEtAL.ma"
METHY="/gpfs/gpfs01/polaris/Q0286/uqywu16/omics/methy/bl_meqtl_cis_std_chr11"
rug="/shares/compbio/Group-Wray/YanHoltz/DATA/EQTL/eQTL_DATA_EQTLGEN_CONSORTIUM_32K/cis-eQTLs-full_formatted.txt_besd-dense"
genelist="/shares/compbio/Group-Wray/YanHoltz/DATA/GENE_POSITION/glist_hg19_refseq.txt"
probe="ENSG00000254682"
window="1000"
psmr="5e-8"
thread="6"
out="./integration_result_Chr11_71Mb"

# Run the analysis
tmp_command="smr_Linux --bfile $ref --gwas-summary $sum --beqtl-summary $METHY  --beqtl-summary $rug --plot --probe $probe --probe-wind $window --gene-list $genelist --psmr $psmr --out $out --thread-num $thread"
qsubshcom "$tmp_command" 1 30G smr_integration 10:00:00 ""

# Transfer locally
cd /Users/y.holtz/Dropbox/QBI/4_UK_BIOBANK_GWAS_PROJECT/VitaminD-GWAS/0_DATA/SMR_PLOT_DATA
scp y.holtz@delta.imb.uq.edu.au:/shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION/plot/* .
```






#- Make the plot {.tabset}
***

How to read the plots (from top to bottom):

- GWAS layer, 
    - **grey dots** = SNPs. Y axis shows their GWAS P-value.
    - **maroon rhombus** = gene expression probe. **navy cycle** = methylation probe. Y axis shows their SMR P-value.
    - **solid** = passed HEIDI threshold. **hollow** = did not pass.
    
- eQTL layers, 
    - One graphic for each probe that passed the **e**SMR and HEIDI tests
    - **maroon cross** = SNPs. The Y axis show their eQTL pvalue = how much this SNP is linked with the gene expression.

- mQTL layers, 
    - One graphic for each probe that passed the **m**SMR and HEIDI tests
    - **blue cycles** = SNPs. The Y axis show their mQTL pvalue = how much this SNP is linked with the gene expression.

- Chromatin layer,
    - Each row of the heatmap show one of the 127 sample from the [Roadmap Epigenomics Mapping Consortium](http://www.roadmapepigenomics.org) (REMC)
    - Each column is a SNP. Its position on the X axis is a proxy of its position on the chromosome.
    - The color of the square gives the function of the SNP, deduced from chromatin state. [Legend detail here](http://egg2.wustl.edu/roadmap/web_portal/imputed.html#chr_imp). Tx = Transcription. Red=promotor.
    - Left color bar: each sample has been observed for a specific tissue. For example, over the 127 samples, about 13 have had measurment in blood & T-cell.
    

```{r, eval=FALSE}
# good folder
cd /shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION

# Create an R script to make the plot for a specific region and put it in a file script.R:
# Read the data file in R:
args <- commandArgs(TRUE)
input <- args[1]
output <- args[2]
source("/shares/compbio/Group-Wray/YanHoltz/SOFT/plot_OmicsSMR_24_April.r")
SMRData = ReadomicSMRData(input)

# Plot
pdf(output, height=15, width=18)
omicSMRLocusPlot(data=SMRData,
                 
                 # Threshold for SMR tests:
                 esmr_thresh=0.1, msmr_thresh=0.1, m2esmr_thresh=0.1,
                 
                 # Threshold for HEIDI tests:
                 esmr_heidi = 0, msmr_heidi = 0, m2esmr_heidi = 0,
                 
                 # Other features
                 max_anno_probe = 6, anno_methyl=TRUE, max_plot_mprobe = 1, window = 700, epi_plot=T )
dev.off()

# Send the script for all the regions I need.
tmp_command="Rscript script.R /shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION/plot/integration_result_Chr11_15Mb.ENSG00000152270.txt plot_integ_ENSG00000152270.pdf"
qsubshcom "$tmp_command" 1 10G smr_plot 10:00:00 ""

tmp_command="Rscript script.R /shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION/plot/integration_result_Chr11_71Mb.ENSG00000254682.txt plot_integ_ENSG00000254682.pdf"
qsubshcom "$tmp_command" 1 10G smr_plot 10:00:00 ""


# Trasfer result locally
cd
scp y.holtz@delta.imb.uq.edu.au://shares/compbio/Group-Wray/YanHoltz/VITAMIND_XIA_ET_AL/10_INTEGRATION/*pdf .
```

## Chromo 11 ~ 15Mb

<img src="IMG/SMR_PLOT/plot_integ_ENSG00000152270.pdf" alt="some text"  width="4200" height="4200">

## Chromo 11 ~ 71 Mb

<img src="IMG/SMR_PLOT/plot_integ_ENSG00000254682.pdf" alt="some text"  width="4200" height="4200">









Detail of function arguments:
```{r, eval=FALSE}
# Plot caractéristics
data=SMRData
esmr_thresh=0.05
msmr_thresh=0.05
m2esmr_thresh=0.05
esmr_heidi = 0
msmr_heidi = 0
m2esmr_heidi = 0
probeNEARBY=c("ENSG00000152270")
interest_only=T
max_anno_probe = 6
anno_methyl=T
max_plot_mprobe = 1
window = 700
epi_plot=T
esmr_thresh_plot=NULL
msmr_thresh_plot=NULL
highlight=NULL
pointsize=20
anno_self=TRUE
ASO=TRUE
anno_dist=1.05
trait_name=NULL
```

