---
title: "A Vitamin-D GWAS in the UK Biobank"
author: "by [Yan Holtz](https://github.com/holtzy/) - `r format(Sys.time(), '%d %B %Y')`"
---


<br><br>


This document summarizes my work concerning the Vitamin-D GWAS in the UKB. It is composed by several parts:

* **Phenotype description**: explains - how to get the Vitamin-D phenotype - clean it - interaction with period. 
* **run GWAS**: the code I used to run the GWAS with plink and description of its result.
* **SMR**: description of the SMR pipeline to find putative genes under the discovered SNPs.
* **GSMR**: description of the GSMR pipeline to find out if Vitamin-D has a causal effect on several mental diseases.
* **XIa et al**: GSMR run on their GWAS summary statistics.


<br><br>

Some of the script described here are run on 2 different clusters:

- Delta (IMB Cluster)
```{r, eval=FALSE}
# Access to delta
ssh y.holtz@delta.imb.uq.edu.au

# delta repository for this work
cd /home/y.holtz/BLOOD_GWAS

# Medici on delta:
cd /gpfs/gpfs01/polaris/Q0286/
```

- Inode (QBI Cluster)
```{r, eval=FALSE}
# Access to Inode
ssh uqyholtz@inode.qbi.uq.edu.au:/

# Inode repo for this work
/ibscratch/wrayvisscher/Yan_Holtz/4_UKB_GWAS/3_BLOOD_GWAS
```
